---
title: "Meta XWAS of HIV+ across sub-Saharan Africa"
date: "06/29/2020"
output: html_document
---


# Authors 
- Chirag J Patel
    + github: [\@chiragjp](http://github.com/chiragjp)
    + web: [www.chiragjpgroup.org](www.chiragjpgroup.org)
- Kajal Claypool
- Eric Chow
- Jake Chung
- Don Mai
- Jessie Chen
- Eran Bendavid
    + email: ebd[at]stanford[dot]edu

# Github Repo and Summary Statistics
- https://github.com/chiragjp/dhs_hiv_meta


```{r, echo=FALSE, warning=FALSE, message=FALSE}

library(tidyverse)
library(scales)
library(ggrepel)
library(DT)
library(gt)
library(gplots)
library(cowplot)
library(plotly)
load('./meta_data/meta_filtered_data.Rdata')
```



```{r, echo=FALSE, include=FALSE}
num_countries_per_var <- meta_country_simple_filtered %>% group_by( name, gender) %>% summarize(num_country=n()) %>% ungroup()
#number_of_associations_country <- meta_country_simple_filtered %>% group_by(k, country, gender) %>% summarize(n=n())
quantile(num_countries_per_var$num_country[num_countries_per_var$gender == 'm'], probs=c(.5, .75, .8, .9, .95, .975, .99))
quantile(num_countries_per_var$num_country[num_countries_per_var$gender == 'f'], probs=c(.5, .75, .8, .9, .95, .975, .99))
num_countries_per_var_to_merge <- num_countries_per_var %>% mutate(num_country_bin = cut(num_country, breaks=c(0,1,10,20,30))) %>% unite(key, name, gender)

#meta_combined_simple_filtered$num_surveys <- cut(meta_combined_simple_filtered$k, breaks=c(0,1,5,10,15,20,25,30, 35,40, 45, 50))
meta_combined_simple_filtered$num_surveys <- cut(meta_combined_simple_filtered$k, breaks=c(0,1,10,20,30,40, 50))
meta_combined_simple_filtered <- meta_combined_simple_filtered %>% unite(key, name, gender, remove = F) %>% left_join(num_countries_per_var_to_merge)

```

# R^2 vs. -log10(pvalue)
- stratified by country number and gender

```{r echo=FALSE}
## how many identified per each bin
R2_thresh <- 0.001 # top 25%
p_thresh <- 1e-6
meta_combined_simple_filtered <- meta_combined_simple_filtered %>% mutate(nlp=-log10(pvalue))
meta_combined_simple_filtered <- meta_combined_simple_filtered %>% mutate(nlp=ifelse(nlp > 100, 100, nlp))

p <- ggplot(meta_combined_simple_filtered, aes(mean_r2, nlp))
p <- p + geom_point(alpha=0.05, color='blue') + scale_x_continuous(limits=c(0, .02)) + scale_y_continuous(limits=c(0, 100))
p <- p + geom_hline(yintercept = -log10(p_thresh), linetype="dashed") + geom_vline(xintercept=R2_thresh, linetype="dashed")
p <- p + facet_grid(num_country_bin ~ gender, labeller=labeller(gender = c(f = "Female", m = "Male"))) 
p <- p + geom_point(alpha=.7, data=meta_combined_simple_filtered %>% filter(pvalue < p_thresh, mean_r2 >= R2_thresh), aes(mean_r2, nlp))
p <- p + xlab('Average R^2 Across Surveys') + ylab('-log10(pvalue)')
p
```

# Odds Ratio vs. -log10(pvalue)
```{r echo=FALSE}
meta_combined_simple_filtered <- meta_combined_simple_filtered %>% mutate(odds_ratio=exp(beta))
meta_combined_simple_filtered <- meta_combined_simple_filtered %>% mutate(odds_ratio2=case_when
                                                                          (odds_ratio > 100 ~ 100,
                                                                            odds_ratio < 0.01 ~ 0.01,
                                                                            TRUE ~ odds_ratio)) 
p <- ggplot(meta_combined_simple_filtered, aes(odds_ratio2, nlp))
p <- p + geom_point(alpha=0.05, color='blue') + scale_y_continuous(limits=c(0, 100)) + scale_x_log10(limits=c(.01,100))
p <- p + geom_hline(yintercept = -log10(p_thresh), linetype="dashed")
p <- p + facet_grid(num_country_bin ~ gender,labeller=labeller(gender = c(f = "Female", m = "Male"))) 
p <- p + geom_point(alpha=.7, data=meta_combined_simple_filtered %>% filter(pvalue < p_thresh, mean_r2 >= R2_thresh), aes(odds_ratio2, nlp))
p <- p + xlab('Odds Ratio Across Surveys') + ylab('-log10(pvalue)')
p
```

# I^2 vs. -log10(pvalue)
```{r echo=FALSE}
p <- ggplot(meta_combined_simple_filtered %>% filter(num_country > 1, pvalue < p_thresh,mean_r2 > R2_thresh), aes(mean_r2, I2))
p <- p + geom_hline(yintercept=50, linetype='dashed')
p <- p + geom_point(alpha=0.7) + scale_x_continuous(limits=c(0, .02)) +facet_grid(num_country_bin ~ gender) 
p <- p + xlab('I^2 of Priortized Factors')
p
```




# Describing the distribution of the associations
## Summary Tables
```{r echo=FALSE}
identified <- meta_combined_simple_filtered %>% mutate(sig=(pvalue < p_thresh) & (mean_r2 >= R2_thresh))
#identified %>% filter(sig == T) %>% group_by(gender) %>% tally()
#identified %>% filter(sig == T) %>% group_by(sig, gender, num_country_bin) %>% tally()

association_summary_table <- identified %>% group_by(num_country_bin, gender) %>% summarize(n=n(),
  q_25_or=quantile(exp(abs(beta)),probs=.25), q_50_or=quantile(exp(abs(beta)),probs=.5), q_75_or = quantile(exp(abs(beta)),probs=.75), q_25_I2=quantile(I2,probs=.25), q_50_I2=quantile(I2,probs=.5),q_75_I2 = quantile(I2,probs=.75),
  q_25_r2=quantile(mean_r2,probs=.25), q_50_r2=quantile(mean_r2,probs=.5), q_75_r2 = quantile(mean_r2,probs=.75)
  )


association_summary_table %>% gt() %>% fmt_number(columns = contains('q_'), decimals = 2) %>% fmt_scientific(columns = ends_with("r2"))

association_summary_table_sig <- identified %>% filter(sig == T) %>% group_by(num_country_bin, gender) %>% summarize(n=n(),
  q_25_or=quantile(exp(abs(beta)),probs=.25), q_50_or=quantile(exp(abs(beta)),probs=.5), q_75_or = quantile(exp(abs(beta)),probs=.75), q_25_I2=quantile(I2,probs=.25), q_50_I2=quantile(I2,probs=.5),q_75_I2 = quantile(I2,probs=.75),
  q_25_r2=quantile(mean_r2,probs=.25), q_50_r2=quantile(mean_r2,probs=.5), q_75_r2 = quantile(mean_r2,probs=.75)
  )

association_summary_table_sig %>% gt() %>% fmt_number(columns = contains('q_'), decimals = 2) %>% fmt_scientific(columns = ends_with("r2"))


association_summary_table_sig_29 <- identified %>% filter(sig == T, num_country==29) %>% group_by(gender) %>% summarize(n=n(),
  q_25_or=quantile(exp(abs(beta)),probs=.25), q_50_or=quantile(exp(abs(beta)),probs=.5), q_75_or = quantile(exp(abs(beta)),probs=.75), q_25_I2=quantile(I2,probs=.25), q_50_I2=quantile(I2,probs=.5),q_75_I2 = quantile(I2,probs=.75),
  q_25_r2=quantile(mean_r2,probs=.25), q_50_r2=quantile(mean_r2,probs=.5), q_75_r2 = quantile(mean_r2,probs=.75)
  ) %>% gt() %>% fmt_number(columns = contains('q_'), decimals = 2) %>% fmt_scientific(columns = ends_with("r2"))

association_summary_table_sig_29

```



## ECDFs of effect sizes
```{r, echo=FALSE, fig.height=9}
p_ecdf_r2 <- ggplot(identified, aes(mean_r2, color=sig)) 
p_ecdf_r2 <- p_ecdf_r2 + stat_ecdf() + scale_x_continuous(limits=c(0, .03))
p_ecdf_r2 <- p_ecdf_r2 + facet_grid(num_country_bin ~ gender, labeller=labeller(gender = c(f = "Female", m = "Male")))
p_ecdf_r2 <- p_ecdf_r2 + theme(legend.position = 'none') + ylab('Percentile') + xlab('R^2')


p_ecdf_or <- ggplot(identified, aes(exp(abs(beta)), color=sig)) 
p_ecdf_or <- p_ecdf_or + stat_ecdf() + scale_x_log10()
p_ecdf_or <- p_ecdf_or + facet_grid(num_country_bin ~ gender, labeller=labeller(gender = c(f = "Female", m = "Male")))
p_ecdf_or <- p_ecdf_or + theme(legend.position = 'none') + ylab('Percentile') + xlab('Odds Ratio (absolute value)')

p_ecdf_i2 <- ggplot(identified %>% filter(num_country > 1), aes(I2, color=sig)) 
p_ecdf_i2 <- p_ecdf_i2 + stat_ecdf() + scale_colour_discrete(name  ="Identified?")
p_ecdf_i2 <- p_ecdf_i2 + facet_grid(num_country_bin ~ gender, labeller=labeller(gender = c(f = "Female", m = "Male"))) 
p_ecdf_i2 <- p_ecdf_i2 + theme(legend.position = 'bottom') + ylab('Percentile') + xlab('I^2')

plot_grid(p_ecdf_r2, p_ecdf_or, p_ecdf_i2, ncol= 1, labels = c('A', 'B', 'C'), rel_heights = c(2,2,2))

```


# Identified Associations 
- R2 > 0.001 and p-value < 1e-6
- SIG==TRUE 

```{r echo=FALSE, warning=FALSE}
## how many identified per each bin
sig_combined_simple <- identified  %>% filter(sig==TRUE) %>% unite(key, name, gender, remove = F)

meta_country_summary <- meta_country_simple_filtered %>% group_by(name, gender) %>% summarize(num_country_2=n(), num_sig_country=sum(pvalue<p_thresh & mean_r2>R2_thresh)) %>% mutate(pct_sig_country=(num_sig_country/num_country_2)*100) %>% ungroup() %>% unite(key, name, gender, remove = T)
sig_combined_simple <- sig_combined_simple %>% left_join(meta_country_summary, by='key')


identified_by_group <- inner_join(
  identified %>% group_by(gender, sig, num_country_bin) %>% tally() %>% ungroup() %>% filter(sig == T) %>% select(-sig) %>% rename(total_sig=n) %>% unite(key, gender, num_country_bin),
  identified %>% group_by(gender, sig, num_country_bin) %>% tally() %>% ungroup() %>% filter(sig == F) %>% select(-sig) %>% rename(total=n) %>% unite(key, gender, num_country_bin),
by="key") %>% separate(key, c("gender", "num_country_bin"), sep="\\_") %>% mutate(pct_sig = total_sig/total)

#identified %>% filter(num_country == 29) %>% group_by(gender, sig) %>% tally() %>% gt()
identified %>% group_by(gender, sig) %>% tally() %>% gt()
```


```{r echo=FALSE, fig.width=11, fig.height=5}
#### preparing data frames for plotting

plot_panels <- function(filtered_frame_of_assoc, top_n_to_plot=10) {
  sample_frame <- filtered_frame_of_assoc
  #sample_frame <- sample_frame %>% arrange(desc(mean_r2)) %>% head(n=top_n_to_plot)
  sample_frame <- sample_frame %>% arrange(desc(pct_sig_country)) %>% head(n=top_n_to_plot)
  sample_frame <- sample_frame %>% arrange(desc(pct_sig_country))
  sample_gender <- sample_frame$gender[1]
  sample_country_frame <- meta_country_simple_filtered %>% filter(gender == sample_gender, name %in% sample_frame$name)
  sample_country_frame <- sample_country_frame %>% unite(key, name, gender, remove = F) %>% left_join(sample_frame %>% select(key, pct_sig_country), by='key')

  sample_frame$name <- reorder(sample_frame$name, sample_frame$pct_sig_country)
  sample_country_frame$name <- reorder(sample_country_frame$name, sample_country_frame$pct_sig_country)
  sample_frame <- sample_frame %>% unite(labelz, name, var_lbl, lvl_lbl, gender, sep="/", remove = F)
  sample_country_frame <- sample_country_frame %>% unite(labelz, name,var_lbl, lvl_lbl, gender, sep="/", remove = F)
  sample_frame$labelz <- reorder(sample_frame$labelz, sample_frame$pct_sig_country)
  sample_country_frame$labelz <- reorder(sample_country_frame$labelz, sample_country_frame$pct_sig_country)

  p1 <- ggplot(sample_frame, aes(labelz, pct_sig_country))
  p1 <- p1 + geom_bar(stat='identity') + ylab('Percent of Countries Variable was Identified') + xlab('') + coord_flip()
  p1 <- p1 + ggtitle('') + theme(plot.title = element_text(size=8), axis.title.x = element_text(size=7)) 

  sample_frame$i2_label <- sprintf('%.1f', sample_frame$I2)
  p3 <- ggplot(sample_frame, aes(labelz, I2))
  p3 <- p3 + geom_bar(stat='identity') + geom_text(aes(label=i2_label)) + ylab('I^2') + xlab('') + coord_flip()
  p3 <- p3 + ggtitle('') + theme(axis.text.y=element_blank(), axis.ticks.y=element_blank(),
                               plot.title = element_text(size=8), axis.title.x = element_text(size=7)) 
  
  p2 <- ggplot(sample_frame, aes(labelz, mean_r2))
  p2 <- p2 + geom_jitter(data=sample_country_frame, aes(labelz, mean_r2, color=sign(beta))) 
  p2 <- p2 + geom_point(color='red', size=3) + scale_y_continuous(limits=c(0, 0.08))
  p2 <- p2 + coord_flip() + xlab('') + ylab("Negelkerke R^2")
  p2 <- p2 + ggtitle('') + theme(axis.text.y=element_blank(), axis.ticks.y=element_blank(),
                               legend.position = 'none', 
                               plot.title = element_text(size=8), axis.title.x = element_text(size=7)) 

  sample_frame$or_label <- sprintf('%.1f', exp(sample_frame$beta))
  p4 <- ggplot(sample_frame, aes(labelz, exp(beta)))
  p4 <- p4 + geom_jitter(data=sample_country_frame, aes(labelz, exp(beta), color=sign(beta)), position = position_jitter(seed=1))
  p4 <- p4 + geom_point(color='red', size=3) 
  p4 <- p4 + geom_errorbar(aes(ymin=exp(beta-2*se), ymax=exp(beta+2*se)),width=.5)
  p4 <- p4 + geom_text_repel(aes(label=or_label))
  p4 <- p4 + scale_y_log10()
  p4 <- p4 + coord_flip() + xlab('') + ylab('Odds Ratio')
  p4 <- p4 + ggtitle('') + theme(axis.text.y=element_blank(), 
                               axis.ticks.y=element_blank(),legend.position = 'none',
                               plot.title = element_text(size=8), axis.title.x = element_text(size=7))
  return(list(or=p4, r2=p2, i2=p3, sig=p1))
}

plot_the_grid <- function(plots) {
  p <- plot_grid(plots$sig,plots$r2,plots$or,plots$i2, nrow=1, rel_widths = c(2,.5,.5,.25))  
}


plots_low_range_f <- plot_panels(sig_combined_simple %>% filter(gender == 'f', num_country > 2, num_country <= 10),10)
plots_mid_range_f <- plot_panels(sig_combined_simple %>% filter(gender == 'f', num_country <= 19, num_country > 10),15)
plots_hi_range_f <- plot_panels(sig_combined_simple %>% filter(gender == 'f', num_country > 19, num_country <= 29), 15)
plots_29_f <- plot_panels(sig_combined_simple %>% filter(gender == 'f', num_country == 29), 15)
#plot_the_grid(plots_low_range_f)
p_mid_f <- plot_the_grid(plots_mid_range_f)
p_hi_f <- plot_the_grid(plots_hi_range_f)
p_29_f <- plot_the_grid(plots_29_f)

```


# Top Associations for females across 20-29 countries in Sub-Saharan Africa
```{r, echo=FALSE, fig.width=12, fig.height=5}
p_hi_f
```
## Table of females across 20-29 countries in Sub-Saharan Africa

```{r, echo=FALSE}
datatable(sig_combined_simple %>% filter(gender == 'f', num_country > 19, num_country <= 29) %>% select(name, var_lbl, lvl_lbl, odds_ratio, nlp, I2, mean_r2, num_country, pct_sig_country)) %>% formatRound('odds_ratio', 3) %>% formatRound('nlp', 2) %>% formatRound('I2', 0) %>% formatRound('mean_r2', 5) %>% formatRound('pct_sig_country',0)
```

```{r echo=FALSE, fig.width=11, fig.height=5}
plots_low_range_m <- plot_panels(sig_combined_simple %>% filter(gender == 'm', num_country > 5, num_country <= 10),10)
plots_mid_range_m <- plot_panels(sig_combined_simple %>% filter(gender == 'm', num_country <= 19, num_country > 10),15)
plots_hi_range_m <- plot_panels(sig_combined_simple %>% filter(gender == 'm', num_country > 19, num_country <= 29), 15)
plots_29_m <- plot_panels(sig_combined_simple %>% filter(gender == 'm', num_country == 29), 15)
#plot_the_grid(plots_low_range_f)
p_mid_m <- plot_the_grid(plots_mid_range_m)
p_hi_m <- plot_the_grid(plots_hi_range_m)
p_29_m <- plot_the_grid(plots_29_m)

```


# Top Associations for males across 20-29 countries in Sub-Saharan Africa
```{r, echo=FALSE, fig.width=12, fig.height=5}
p_hi_m
```

## Table of males across 20-29 countries in Sub-Saharan Africa
```{r, echo=FALSE}
datatable(sig_combined_simple %>% filter(gender == 'm', num_country > 19, num_country <= 29) %>% select(name, var_lbl, lvl_lbl, odds_ratio, nlp, I2, mean_r2, num_country,pct_sig_country)) %>% formatRound('odds_ratio', 3) %>% formatRound('nlp', 2) %>% formatRound('I2', 0) %>% formatRound('mean_r2', 5) %>% formatRound('pct_sig_country',0)
```

# Concordance of Odds Ratios of males vs. females across Sub-Saharan Africa
- comparing the associations across countries by correlating their odds ratios in males versus females
- association has to occur in at least 2 countries
- Those labeled "Female" are identified in females (and not in males); those labeled with "Male" are those identified in males
- Those labeled "Male+Female" are identified in both males and females

```{r, echo=FALSE}
females <- identified %>% filter(gender == 'f', num_country > 1)
males <- identified %>% filter(gender == 'm', num_country > 1)
wide_f_m <- females %>% inner_join(males, by='name', suffix=c('_female', '_male'))
wide_f_m <- wide_f_m %>% mutate(identified_in=case_when(
  sig_female & sig_male == TRUE ~ "Male+Female",
  sig_female == TRUE ~ "Female",
  sig_male == TRUE ~ "Male",
  TRUE ~ 'Neither'
))
p <- ggplot(wide_f_m, aes(exp(beta_female), exp(beta_male)))
p <- p + geom_point(alpha=0.5) + scale_y_log10(limits=c(0.001, 100)) + scale_x_log10(limits=c(0.001, 100)) + facet_wrap(~identified_in, ncol=2)
p <- p + xlab('Odds Ratios [female]') + ylab("Odds Ratio [male]") + geom_abline()
ggplotly(p)

#wide_f_m %>% group_by(identified_in) %>% summarize(n=n(), cr = cor(beta_female, beta_male))

```


# Country by country correlation for males
- correlate the associations that are assessed in all 29 countries between each of the countries

```{r echo=FALSE, fig.height=12, fig.width=12}

heatmapColors <- function(numColors=16) {
	c1 <- rainbow(numColors,v=seq(0.5,1,length=numColors),s=seq(1,0.3,length=numColors),start=4/6,end=4.0001/6);
	c2 <- rainbow(numColors,v=seq(0.5,1,length=numColors),s=seq(1,0.3,length=numColors),start=1/6,end=1.0001/6);
	c3 <- c(c1,rev(c2)); 
	return(c3)
}

vars_in_all_countries <- num_countries_per_var %>% filter(num_country == 29, gender=='m')
meta_country_all_m <- meta_country_simple_filtered %>% filter(name %in% vars_in_all_countries$name)  %>% filter(gender=='m') %>% select(beta, name, country) 
meta_country_all_m_wide <- meta_country_all_m %>% spread(name, beta)

meta_country_all_m_wide_matr <- as.matrix(meta_country_all_m_wide[, 2:ncol(meta_country_all_m_wide)])
rownames(meta_country_all_m_wide_matr) <- meta_country_all_m_wide[,1]

cr_m <- cor(t(meta_country_all_m_wide_matr), use='pairwise.complete.obs', method = 'spearman')
quantile(abs(cr_m[upper.tri(cr_m)]), probs=c(0.25, 0.5, 0.75))
heatmap.2(cr_m, trace = 'none', na.rm = F, scale = 'none', symm = T, col=heatmapColors(5), margins=c(16,16), sepwidth=c(.1, .1), symbreaks=T)

```

# Country by country correlation for females
```{r echo=FALSE, fig.height=12, fig.width=12}
vars_in_all_countries <- num_countries_per_var %>% filter(num_country == 29,  gender=='f')
meta_country_all_f <- meta_country_simple_filtered %>% filter(name %in% vars_in_all_countries$name)  %>% filter(gender=='f') %>% select(beta, name, country) 
meta_country_all_f_wide <- meta_country_all_f %>% spread(name, beta)


meta_country_all_f_wide_matr <- as.matrix(meta_country_all_f_wide[, 2:ncol(meta_country_all_f_wide)])
rownames(meta_country_all_f_wide_matr) <- meta_country_all_f_wide[,1]

cr_f <- cor(t(meta_country_all_f_wide_matr), use='pairwise.complete.obs', method = 'spearman')
quantile(abs(cr_f[upper.tri(cr_f)]), probs=c(0.25, 0.5, 0.75))

heatmap.2(cr_f, trace = 'none', na.rm = F, scale = 'none', symm = T, col=heatmapColors(5), margins=c(16,16), sepwidth=c(.1, .1), symbreaks=T)
```

