---
title: "Figures and Tables for Meta XWAS in HIV"
author: "Chirag Patel"
date: "8/27/2019"
output: html_document
---



```{r, echo=FALSE, warning=FALSE, message=FALSE}
library(gplots)
library(cowplot)
library(tidyverse)
source('0_load_and_filter_rmd.R')
```
# How many variables interrogated?
```{r}
meta_combined_simple_filtered %>% group_by(gender) %>% summarize(n=n())
print(length(unique(meta_combined_simple_filtered$name)))

```
# quantiles of R2
```{r}
p <- ggplot(meta_combined_simple_filtered, aes(mean_r2))
p <- p + stat_ecdf()
p
p <- ggplot(meta_country_simple_filtered, aes(mean_r2))
p <- p + stat_ecdf()
p
quantile(meta_combined_simple_filtered$mean_r2, probs=c(0.5, .99), na.rm = T)
quantile(meta_country_simple_filtered$mean_r2, probs=c(0.5, .99), na.rm = T)
```

## Distribution of number of surveys and countries a variable appears in
```{r, echo=FALSE, warning=FALSE}
num_vars_by_k <- meta_combined_simple_filtered %>% group_by(k, gender) %>% summarize(num_variables=n())
num_vars_by_k_m <- num_vars_by_k %>% ungroup() %>% filter(gender == 'm')
num_vars_by_k_m[order(num_vars_by_k_m$k),'cumsum'] <- cumsum(num_vars_by_k_m[order(num_vars_by_k_m$k),'num_variables'] )

num_vars_by_k_f <- num_vars_by_k %>% ungroup() %>% filter(gender == 'f')
num_vars_by_k_f[order(num_vars_by_k_f$k),'cumsum'] <- cumsum(num_vars_by_k_f[order(num_vars_by_k_f$k),'num_variables'] )
num_vars_by_k <- rbind(num_vars_by_k_m, num_vars_by_k_f)

p <- ggplot(num_vars_by_k , aes(k, cumsum))
p <- p + geom_point() + geom_line() + facet_wrap(~gender, nrow=1) + ylab('Cumulative Number of Variables') + xlab('Number of surveys')
p

num_countries_per_var <- meta_country_simple_filtered %>% group_by( name, gender) %>% summarize(num_country=n()) %>% ungroup()

country_var_g <- num_countries_per_var %>% group_by(gender, num_country) %>% summarize(num_variables=n()) %>% ungroup()
num_vars_by_country_m <- country_var_g %>% filter(gender == 'm')
num_vars_by_country_m[order(num_vars_by_country_m$num_country),'cumsum'] <- cumsum(num_vars_by_country_m[order(num_vars_by_country_m$num_country),'num_variables'] )
num_vars_by_country_f <- country_var_g %>%  filter(gender == 'f')
num_vars_by_country_f[order(num_vars_by_country_f$num_country),'cumsum'] <- cumsum(num_vars_by_country_f[order(num_vars_by_country_f$num_country),'num_variables'] )
num_vars_by_country <- rbind(num_vars_by_country_m, num_vars_by_country_f)

p <- ggplot(num_vars_by_country , aes(num_country, cumsum))
p <- p + geom_point() + geom_line() + facet_wrap(~gender, nrow=1) + ylab('Cumulative Number of Variables') + xlab('Number of Countries')
p

p <- ggplot(num_vars_by_k, aes(k, num_variables)) 
p <- p + geom_bar(stat = 'identity') + xlab('Number of Surveys') + ylab('Number of Variables')
p <- p + facet_wrap(~gender, nrow=1) + scale_y_log10()
p

p <- ggplot(num_vars_by_country, aes(num_country, num_variables)) 
p <- p + geom_bar(stat = 'identity') + xlab('Number of Countries') + ylab('Number of Variables')
p <- p + facet_wrap(~gender, nrow=1) + scale_y_log10()
p
```

# What are the percentiles of the distribution of number of variables
```{r}
quantile(num_countries_per_var$num_country[num_countries_per_var$gender == 'm'], probs=c(.5, .75, .8, .9, .95, .975, .99))
quantile(num_countries_per_var$num_country[num_countries_per_var$gender == 'f'], probs=c(.5, .75, .8, .9, .95, .975, .99))
num_countries_per_var_to_merge <- num_countries_per_var %>% mutate(num_country_bin = cut(num_country, breaks=c(0,1,5,10,15,20,25,30))) %>% unite(key, name, gender)
table(num_countries_per_var_to_merge$num_country_bin)
```





```{r echo=FALSE, warning=FALSE}
## how many identified per each bin
R2_thresh <- 0.01 
identified <- meta_country_simple_filtered %>% mutate(sig=(pvalue < 1e-6) & (mean_r2 >= R2_thresh))

sig_country_simple <- meta_country_simple_filtered %>% mutate(sig=(pvalue < 1e-6) & (mean_r2 >= R2_thresh)) %>% filter(sig==TRUE) %>% unite(key, name, gender, remove = F)

identified <- identified %>% group_by(name, gender) %>% summarise(n_sig=sum(sig), total=n()) %>% mutate(pct_sig=100*(n_sig/total)) %>% ungroup() #%>% #
identified <- left_join(identified %>% unite(key, name, gender, remove=F, sep="/"), meta_combined_simple_filtered %>% unite(key,name, gender,sep='/') %>% select(key, beta, se, I2, mean_r2) %>% transmute(key=key,overall_mean_r2=mean_r2, overall_beta=beta, overall_se=se, overall_I2=I2), by='key') %>% arrange(desc(pct_sig)) %>% mutate(key=NULL)

var_labels <- meta_country_simple_filtered %>% select(name, gender, var_lbl, lvl_lbl) %>% unique() %>% unite(label, var_lbl, lvl_lbl, gender, sep="/", remove = F)
var_labels$var_lbl <- NULL
var_labels$lvl_lbl <- NULL

```

## Top N associations identified by:
- Significance (Bonferroni signficance and mean_r2 over all countries > 0.01)
- number of countries a variable was signficant (rule 1)
- create a list for Kajal

```{r echo=FALSE, warning=FALSE}

NUMBER_TO_SHOW <- 20
identified_bins <- identified %>% unite(key, name, gender, remove = F) %>% left_join(num_countries_per_var_to_merge, by='key') 
identified_bins <- identified_bins %>% left_join(var_labels %>% unite(key, name, gender), by='key')
identified_bins <- identified_bins %>% filter(n_sig > 0)
identified_bins_plot <- identified_bins %>% filter(num_country >5 ) %>% group_by(num_country_bin, gender) %>% group_split() 
identified_bins_plot <- identified_bins_plot %>% map(~arrange(., desc(.$n_sig))) %>% map(~head(., n=NUMBER_TO_SHOW))

## plot for males and females and bar chart
b_pct_plot <- function(frm) {
  main_title <- sprintf('Number of Countries:%s; Gender:%s',frm$num_country_bin[1], frm$gender[1])
  frm$label <- paste(frm$name, frm$label)
  frm$label <- factor(frm$label, levels = frm$label[order(frm$pct_sig, decreasing = F)])
  p <- ggplot(frm, aes(label, pct_sig)) + geom_bar(stat='identity') + ylab('Percent of Countries Variable was Identified') +xlab('') + coord_flip()
  p <- p + ggtitle(main_title)+theme(plot.title = element_text(size=8), axis.title.x = element_text(size=7)) 
  p
}

identified_bins_pct_plot_obj <- identified_bins_plot %>% map(~b_pct_plot(.))

```




### Summary Statistics Plot Preparation
```{r echo=FALSE, warning=FALSE}
r2_scatter_plot <- function(frm) {
  main_title <- sprintf('Number of Countries:%s; Gender:%s',frm$num_country_bin[1], frm$gender[1])
  p <- ggplot(frm, aes(label, mean_r2, col=sign(beta)))
  p <- p + geom_jitter() 
  p <- p + ylab('Country-level R2') + xlab('') 
  p <- p + geom_point(data=frm,mapping=aes(label, overall_mean_r2), )
  p <- p + coord_flip()
  p <- p + ggtitle(main_title) + theme(legend.position = 'none',plot.title = element_text(size=8), axis.title.x = element_text(size=7))
  p
}

log_or_scatter_plot <- function(frm) {
  main_title <- sprintf('Number of Countries:%s; Gender:%s',frm$num_country_bin[1], frm$gender[1])
  p <- ggplot(frm, aes(label, beta, col=sign(beta)))
  p <- p + geom_jitter() 
  p <- p + ylab('Country-level log(OR)') + xlab('') + coord_flip()
  p <- p + ggtitle(main_title) + theme(legend.position = 'none',plot.title = element_text(size=8), axis.title.x = element_text(size=7))
  p
}


r2_plot_objs <- vector('list', length(identified_bins_plot))
or_plot_objs <- vector('list', length(identified_bins_plot))
for(index in seq_along(identified_bins_plot)) {
  plot_data <- identified_bins_plot[[index]]
  plot_data$label <- paste(plot_data$name, plot_data$label)
  plot_data$label <- factor(plot_data$label, levels = plot_data$label[order(plot_data$pct_sig, decreasing = F)])
  r2_plot_objs[[index]] <- r2_scatter_plot(plot_data %>% left_join(sig_country_simple %>% select(key,country, beta, mean_r2, pvalue), by='key'))
  or_plot_objs[[index]] <- log_or_scatter_plot(plot_data %>% left_join(sig_country_simple %>% select(key,country, beta, mean_r2, pvalue), by='key'))
}



```

## Display in a paneled grid
```{r echo=FALSE, fig.height=6, fig.width=18, warning=FALSE}
for(index in seq_along(identified_bins_plot)) {
  pg <- plot_grid(identified_bins_pct_plot_obj[[index]], 
          r2_plot_objs[[index]] + theme(axis.text.y=element_blank(), axis.ticks.y=element_blank()) + ggtitle(''), 
          or_plot_objs[[index]] + theme(axis.text.y=element_blank(), axis.ticks.y=element_blank()) + ggtitle('') + scale_y_continuous(limits=c(-4,4)), 
          labels = c('A', 'B', 'C'), label_size = 12, nrow = 1, rel_widths = c(2, 1,1))
  print(pg)
}


```


## Plot the entire distribution
```{r echo=FALSE, warning=FALSE}

r2_scatter_plot_all <- function(frm) {
  main_title <- sprintf('Number of Countries:%s; Gender:%s',frm$num_country_bin[1], frm$gender[1])
  p <- ggplot(frm %>% filter(sig==TRUE), aes(label, mean_r2, col=sign(beta)))
  p <- p + geom_point() 
  p <- p + geom_point(data=frm %>% filter(sig==FALSE), mapping=aes(label, beta),alpha=.5, color='gray')
  p <- p + ylab('Country-level R2') + xlab('') + coord_flip()
  p <- p + ggtitle(main_title) + theme(legend.position = 'none',plot.title = element_text(size=8), axis.title.x = element_text(size=7))
  p
}

log_or_scatter_plot_all <- function(frm) {
  main_title <- sprintf('Number of Countries:%s; Gender:%s',frm$num_country_bin[1], frm$gender[1])
  p <- ggplot(frm %>% filter(sig==TRUE), aes(label, beta, col=sign(beta)))
  p <- p + geom_point()  
  p <- p + geom_point(data=frm %>% filter(sig==FALSE), mapping=aes(label, beta),alpha=.5, color='gray')
  p <- p + ylab('Country-level log(OR)') + xlab('') + coord_flip()
  p <- p + ggtitle(main_title) + theme(legend.position = 'none',plot.title = element_text(size=8), axis.title.x = element_text(size=7))
  p
}

all_country_simple <- meta_country_simple_filtered %>% mutate(sig=(pvalue < 1e-7) & (mean_r2 >= R2_thresh)) %>% unite(key, name, gender)
r2_plot_objs_all <- vector('list', length(identified_bins_plot))
or_plot_objs_all <- vector('list', length(identified_bins_plot))
for(index in seq_along(identified_bins_plot)) {
  plot_data <- identified_bins_plot[[index]]
  plot_data$label <- paste(plot_data$name, plot_data$label)
  plot_data$label <- factor(plot_data$label, levels = plot_data$label[order(plot_data$pct_sig, decreasing = F)])
  r2_plot_objs_all[[index]] <- r2_scatter_plot_all(plot_data %>% left_join(all_country_simple %>% select(key,country, beta, mean_r2, pvalue, sig), by='key'))
  or_plot_objs_all[[index]] <- log_or_scatter_plot_all(plot_data %>% left_join(all_country_simple %>% select(key,country, beta, mean_r2, pvalue,sig), by='key'))
}


```


```{r fig.height=6, fig.width=18 , warning=FALSE}
for(index in seq_along(identified_bins_plot)) {
pg2 <- plot_grid(identified_bins_pct_plot_obj [[index]], 
          r2_plot_objs_all[[index]] + theme(axis.text.y=element_blank(), axis.ticks.y=element_blank()) + ggtitle('') + scale_y_continuous(limits=c(0, .15)), 
          or_plot_objs_all[[index]] + theme(axis.text.y=element_blank(), axis.ticks.y=element_blank()) + ggtitle('') + scale_y_continuous(limits=c(-4,4)), 
          labels = c('A', 'B', 'C'), label_size = 12, nrow = 1, rel_widths = c(2, 1,1))
print(pg2)
}
```

## Tables - one for the prediction and mapping tasks (for Kajal and Jake)
```{r, echo=FALSE}
names_in_29_males <- num_countries_per_var %>% filter(num_country == 29, gender=='m') %>% select(name)
names_in_29_females <- num_countries_per_var %>% filter(num_country == 29, gender=='f') %>% select(name)
sig_29_males <- sig_country_simple %>% filter(gender == 'm', !is.na(beta)) %>% right_join(names_in_29_males, by='name') 
sig_29_females <- sig_country_simple %>% filter(gender == 'f', !is.na(beta)) %>% right_join(names_in_29_females, by='name')
sig_29_countries_males <- meta_combined_simple_filtered %>% filter(gender == 'm') %>% filter( name %in% sig_29_males$name) %>% arrange(desc(mean_r2)) 
sig_29_countries_females <- meta_combined_simple_filtered %>% filter(gender == 'f') %>% filter( name %in% sig_29_females$name) %>% arrange(desc(mean_r2)) 
#write_csv(sig_29_countries_males,path='../dhs_hiv_meta/meta_data/tables_082119/sig_29_countries_males.csv')
#write_csv(sig_29_countries_females,path='../dhs_hiv_meta/meta_data/tables_082119/sig_29_countries_females.csv')

write_csv(sig_29_countries_males,path='../dhs_hiv_meta/meta_data/tables_082119/sig_29_countries_males.csv')
write_csv(sig_29_countries_females,path='../dhs_hiv_meta/meta_data/tables_082119/sig_29_countries_females.csv')


```





# Country by country correlation for females
```{r fig.height=12, fig.width=12 , warning=FALSE}
vars_in_all_countries <- num_countries_per_var %>% filter(num_country == 29,  gender=='f')
meta_country_all_f <- meta_country_simple_filtered %>% filter(name %in% vars_in_all_countries$name)  %>% filter(gender=='f') %>% select(beta, name, country) 
meta_country_all_f_wide <- meta_country_all_f %>% spread(name, beta)


meta_country_all_f_wide_matr <- as.matrix(meta_country_all_f_wide[, 2:ncol(meta_country_all_f_wide)])
rownames(meta_country_all_f_wide_matr) <- meta_country_all_f_wide[,1]

cr_f <- cor(t(meta_country_all_f_wide_matr), use='pairwise.complete.obs', method = 'spearman')
heatmapColors <- function(numColors=16) {
	c1 <- rainbow(numColors,v=seq(0.5,1,length=numColors),s=seq(1,0.3,length=numColors),start=4/6,end=4.0001/6);
	c2 <- rainbow(numColors,v=seq(0.5,1,length=numColors),s=seq(1,0.3,length=numColors),start=1/6,end=1.0001/6);
	c3 <- c(c1,rev(c2)); 
	return(c3)
}

heatmap.2(cr_f, trace = 'none', na.rm = F, scale = 'none', symm = T, col=heatmapColors(5), margins=c(16,16), sepwidth=c(.1, .1), symbreaks=T)

```


# Country by country correlation for males
```{r fig.height=12, fig.width=12}
vars_in_all_countries <- num_countries_per_var %>% filter(num_country == 29, gender=='m')
meta_country_all_m <- meta_country_simple_filtered %>% filter(name %in% vars_in_all_countries$name)  %>% filter(gender=='m') %>% select(beta, name, country) 
meta_country_all_m_wide <- meta_country_all_m %>% spread(name, beta)

meta_country_all_m_wide_matr <- as.matrix(meta_country_all_m_wide[, 2:ncol(meta_country_all_m_wide)])
rownames(meta_country_all_m_wide_matr) <- meta_country_all_m_wide[,1]

cr_m <- cor(t(meta_country_all_m_wide_matr), use='pairwise.complete.obs', method = 'spearman')

heatmap.2(cr_m, trace = 'none', na.rm = F, scale = 'none', symm = T, col=heatmapColors(5), margins=c(16,16), sepwidth=c(.1, .1), symbreaks=T)

```
# Country correlation of correlations
```{r}
cr_f <- cr_f[sort(colnames(cr_f)), sort(colnames(cr_f))]
cr_m <- cr_m[sort(colnames(cr_f)), sort(colnames(cr_f))]
corrcorr_table <- tibble(correlation_females=cr_f[upper.tri(cr_f)], correlation_males=cr_m[upper.tri(cr_m)])
cor(corrcorr_table)
p <- ggplot(corrcorr_table, aes(correlation_females, correlation_males))
p <- p + geom_point()
p <- p + geom_abline()
p
```




